### 2021-04-21 input/proposal.tex
on implem and complexity wip

\noindent This section starts by introducing consistent partitioning
and its implementation when a \process can only become a new source to
create a new partition, i.e., it cannot revoke its self-appointed
status of source. Then, it extends to include the latter behavior: it
introduces the properties that guarantee consistent partitioning in
such context, and it details our implementation called \NAME. Finally,
it provides the complexity analysis of \NAME divided in time, space,
and communication.

@@

One could guarantee consistent partitioning by always propagating the
$\delta$ messages corresponding to the $\alpha$ messages it propagated
before. In Figure~\ref{fig:problem}, it means that as soon as
\Process~$b$ forwards $\alpha_a^2$, it assumes that its neighbors
\Process~$c$ and \Process~$d$ may need the notification of removal
$\delta_a$ if it exists. However, such solution also implies that
\processes generate traffic not only related to their current
partition, but also related to partitions they belong to in the
past. This would prove overly expensive in dynamic systems where
\processes join or leave the system, create or delete partitions, at
any time. Instead, we propose to use the delivery order at each
\process to detect possible inconsistencies and solve them. Together,
\processes eventually remove all stale control information (transiting
messages and local states) of the system leaving room for propagation
of messages about up-to-date partitions.

%% To identify and solve the issue with dynamic consistent partitioning
%% where protocols enable \processes to \texttt{Add} and \texttt{Del} at
%% any time, we need to further constrain the delivery order at each
%% \process.

@@

To implement last win order as stated in Definition~\ref{def:lwo},
each \process maintains a vector of versions that associates the
respective known local counter of each known source, or has-been
source.  Each \process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector enables both Propagation~I and Propagation~I'.  The size
of such vector increases linearly with the number of sources that the
\process ever known which is the number of \processes in the system
$\mathcal{O}(V)$ in the worst case. Unfortunately, \processes cannot
remove any entry from this vector without running a costly distributed
consensus protocol to ensure that no \process can mistake an old but
removed entry for a new one. However, using \NAMEB, we expect that
every \process knows only a small subset of sources.

This vector of versions constitutes a summary of deliveries. \NAME
uses it to detect possible inconsistent partitioning as depicted in
Figure~\ref{fig:proofB}. A detecting \process then propagates a
$\delta$ message conveying the unique identifier $s$ and version $c$
of the partition along with a path $\pi$ comprising the identity of
\processes that forwarded the corresponding $\alpha$ message. Such
$\delta$ message implements the uncertain $\delta?$ message of
Figure~\ref{fig:proofB}. It fulfils the double role of removing stale
control information without modifying the local counter of the source
(since the partition may still exist), and removing loops to ensure
termination (since \processes do not modify their local state to
reflect such delivery, messages piggyback monotonically such
increasing control information). Thus, $\delta$ messages used in
Propagation~I' are more efficient both in propagation time and
traffic, for they propagate using every communication link available
in the partition, and they do not need to carry their path.
Fortunately, Propagation~I' happens more often than Propagation~IV.
Yet, \processes that belong to a partition need to save the path taken
by the corresponding $\alpha$ messages to enable Propagation~IV.  In
the worst case, the path includes every \processes in the system but
converges in average to the average diameter of partitions. Thus,
message paths and \NAME have an interesting synergy: Paths tend to be
smaller as the number of sources in the system increases.

As stated in Theorem~\ref{theo:dcp}, dynamic consistent partitioning
not only requires eventual purging of stale control information, but
also the retrieval of the best up-to-date partitions. In that regard,
$\delta$ messages have dual use: since they already reach the borders
of their partition when they remove stale control information, they
also trigger a competition when reaching such bordering \processes
(see Line~\ref{line:ascast_compete}). This simply consists in sending
the current partition through the communication link from which the
\process received the $\delta$. Upon receipt of this answer,
\processes act normally by propagating their changes when they
improve.

\TODO{About communication complexity} In terms of number of messages
required to reach optimal partitioning. In the average-case, a
\process $i$ chosen uniformly at random among all \processes creates a
logical partition. Its messages $\alpha_i$ propagate through the
network until reaching \processes that belong to another partition
closer to them. This splits partitions in half in average. Overall,
the $a^{th}$ new partition comprises
\smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor \log_2 a \rfloor}})$}
\processes. This decreases every new partition until reaching $0$
\processes per new partition: even the chosen \process already belongs
to its optimal partition. The average number of messages per \process
is \smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 a
      \rfloor}})$}. \TODO{Multiple receipt and multiple delivery imply
  more messages (receipt bounded by $|O|$ as well).} Deleting the
$a^{th}$ partition generates the exact same number of messages than
the $a^{th}$ partition creation. \TODO{But what about echos?} In the
worst-case, every new partition includes all but one \process
belonging to the previous partition. The total number of messages
after the $a^{th}$ new partition is $\mathcal{O}(\overline{|O|}\cdot
a^2)$. As for the average-case, the number of messages for the
partition deletion is identical to the number of messages of the
corresponding partition creation.
\noindent \TODO{In terms of size of messages.}

>>>>

\noindent This section starts by introducing consistent partitioning
and its implementation when a \process can only become a new source to
create a new partition, i.e., it cannot revoke its self-appointed
status of source. Then, it extends to include the latter behavior: it
introduces the properties that guarantee consistent partitioning in
such context, and it details our implementation called \NAME along
with its complexity analysis.

@@

One could guarantee consistent partitioning by always propagating the
$\delta$ messages corresponding to the $\alpha$ messages it propagated
before. In Figure~\ref{fig:problem}, it means that as soon as
\Process~$b$ forwards $\alpha_a^2$, it assumes that its neighbors
\Process~$c$ and \Process~$d$ may need the notification of removal
$\delta_a$ if it exists. However, such solution also implies that
\processes generate traffic not only related to their current
partition, but also related to partitions they belong to in the
past. This would prove overly expensive in dynamic systems where
\processes join or leave the system, create or delete partitions, at
any time.  \TODO{Another paragraph with JJ approach?}  Instead, we
propose to use the delivery order at each \process to detect possible
inconsistencies and solve them. Together, \processes eventually remove
all stale control information (transiting messages and local states)
of the system leaving room for propagation of messages about
up-to-date partitions.

@@

To implement last win order as stated in Definition~\ref{def:lwo},
each \process maintains a vector of versions that associates the
respective known local counter of each known source, or has-been
source.  Each \process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector enables both Propagation~I and Propagation~I'.  The size
of such vector increases linearly with the number of sources that the
\process ever known which is the number of \processes in the system
$\mathcal{O}(V)$ in the worst case. Unfortunately, \processes cannot
remove any entry from this vector without running a costly distributed
consensus protocol to ensure that no \process can mistake an old but
removed entry for a new one. However, following the principles of
scoped broadcast, we expect that every \process knows only a small
subset of sources.

This vector of versions constitutes a summary of deliveries. \NAME
uses it to detect possible inconsistent partitioning as depicted in
Figure~\ref{fig:proofB}. A detecting \process then propagates a
$\delta$ message conveying the unique identifier $s$ and version $c$
of the partition along with a path $\pi$ comprising the identity of
\processes that forwarded the corresponding $\alpha$ message. Such
$\delta$ message implements the uncertain $\delta?$ message of
Figure~\ref{fig:proofB}. It fulfils the double role of removing stale
control information without modifying the local counter of the source
(since the partition may still exist), and removing loops to ensure
termination (since \processes do not modify their local state to
reflect such delivery, messages piggyback monotonically such
increasing control information). Thus, $\delta$ messages used in
Propagation~I' are more efficient both in propagation time and
traffic, for they propagate using every communication link available
in the partition, and they do not need to carry their path.
Fortunately, Propagation~I' happens more often than Propagation~IV.
Yet, \processes that belong to a partition need to save the path taken
by the corresponding $\alpha$ messages to enable Propagation~IV.  In
the worst case, the path includes every \processes in the system but
converges in average to the average diameter of partitions. Thus,
message paths and \NAME have an interesting synergy: Paths tend to be
smaller as the number of sources in the system increases. Bloom
filters could introduce another trade-off by further decreasing the
size of path data structures at the cost of premature stopped
propagation that could impair consistent
partitioning~\cite{whitaker2002forwarding}.

As stated in Theorem~\ref{theo:dcp}, dynamic consistent partitioning
not only requires eventual purging of stale control information, but
also the retrieval of the best up-to-date partitions. In that regard,
both kinds of $\delta$ messages have dual use: since they already
reach the borders of their partition when they remove stale control
information, they also trigger a competition when reaching such
bordering \processes (see Line~\ref{line:ascast_compete}). This simply
consists in sending the current partition through the communication
link from which the \process received the $\delta$. Upon receipt of
this answer, \processes act normally by propagating their changes when
they improve. \TODO{Exemple with figure for detection?}

\TODO{Rework this.}\NAME guarantees dynamic consistent partitioning by
making extensive use of \process to \process communications. In terms
of number of messages, in the average-case, a \process $i$ chosen
uniformly at random among all \processes creates a logical
partition. Its messages $\alpha_i$ propagate through the network until
reaching \processes that belong to another partition closer to
them. This splits partitions in half in average. Overall, the $a^{th}$
new partition comprises \smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor
      \log_2 a \rfloor}})$} \processes. This decreases every new
partition until reaching $0$ \processes per new partition: even the
chosen \process already belongs to its optimal partition. The average
number of messages per \process is
\smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 a
      \rfloor}})$}. \TODO{Multiple receipt and multiple delivery imply
  more messages (receipt bounded by $|O|$ as well).} Deleting the
$a^{th}$ partition generates the exact same number of messages than
the $a^{th}$ partition creation. \TODO{But what about echos?} In the
worst-case, every new partition includes all but one \process
belonging to the previous partition. The total number of messages
after the $a^{th}$ new partition is $\mathcal{O}(\overline{|O|}\cdot
a^2)$. As for the average-case, the number of messages for the
partition deletion is identical to the number of messages of the
corresponding partition creation.
